---
title: "lamda-alpha-estimates"
author: "Jens Johnson"
date: "October 2, 2017"
output: pdf_document
---
# this code generates coexistence parameters for Plectritis and Valerianella 
# from experimentally obtained measures of vital rates

# load libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(purrr)
library(repurrrsive)
library(listviewer)
library(gnlm)
library(boot)
library(propagate)
```

# load data
```{r}
## the Seed_Counts csv contains info on the seed production and seed characteristics of plants grown against different community densities and in a well-watered or drought-stressed scenario.
list.files()
SEEDS <- read.csv("Seed_Counts.csv")
SEEDS <- SEEDS %>% drop_na(num_seeds)


## germination values are assumed to be global, not affected by treatment or seeding densities
germination_i = 0.670
germination_j = 0.891
```

#### Estimate parameters in dry treatment conditions
# aii
```{r}
## Dry
## i = PLCO, j = VALO
##aii
subset_SEEDS <- filter(SEEDS, focal_species == "PLCO", background_species == "PLCO" | background_species == "none", treatment == "Drought-stressed")

#m2_PLCO_aii_drought <- nls(num_seeds~lambda/(1+alpha*germination_i*background_density), data=subset_SEEDS, start=list(lambda=380, alpha=0.22))
#y <- subset_SEEDS$num_seeds*germination_i
#x <- subset_SEEDS$background_density

attach(subset_SEEDS)
mu <- function(p) p[1]/(1+p[2]*background_density)
m2_PLCO_aii_drought <- gnlr(num_seeds, dist="negative binomial", mu=mu, pmu=c(400,0.5), pshape=1)

# Bootstrap 

# function to setimate parameters from given seed_count and background density
mu <- function(p) p[1]/(1+p[2]*background_density)

# function to obtain p1 from the data
p1 <- function(formula, data, indices) {
  d <- sample_n(data, 35) # allows boot to select sample
  attach(d)
  fit <- gnlr(num_seeds, dist="negative binomial", mu=formula, pmu=c(400,0.50), pshape=1)
  return(fit[[18]][1])
}

# bootstrapping with 1000 replications
set.seed(1)
results_lambda_i_dry <- boot(data=subset_SEEDS, p1,
   R=50, formula=mu)

# view results
results_lambda_i_dry


# function to obtain p2 from the data
p2 <- function(formula, data, indices) {
  d <- sample_n(data, 35) # allows boot to select sample
  attach(d)
  fit <- gnlr(num_seeds, dist="negative binomial", mu=formula, pmu=c(400,0.50), pshape=1)
  return(fit[[18]][2])
}

# bootstrapping with 1000 replications
set.seed(1)
results_alpha_ii_dry <- boot(data=subset_SEEDS, p2,
   R=50, formula=mu)

# view results
results_alpha_ii_dry
```

# aij
```{r}
## Dry
## i = PLCO, j = VALO
##aij
subset_SEEDS <- filter(SEEDS, focal_species == "PLCO", background_species == "VALO" | background_species == "none", treatment == "Drought-stressed")

#m2_PLCO_aij_drought <- nls(num_seeds~lambda/(1+alpha*germination_j*background_density), data=subset_SEEDS, start=list(lambda=420, alpha=0.294))
#y<-subset_SEEDS$num_seeds
#x<-subset_SEEDS$background_density

#summary(m2_PLCO_aij_drought)
#confint(m2_PLCO_aij_drought)

attach(subset_SEEDS)
mu <- function(p) p[1]/(1+p[2]*background_density)
m2_PLCO_aij_drought <- gnlr(num_seeds, dist="negative binomial", mu=mu, pmu=c(360,0.30), pshape=1)

# Bootstrap 

# function to setimate parameters from given seed_count and background density
mu <- function(p) p[1]/(1+p[2]*background_density)


# function to obtain p2 from the data
p2 <- function(formula, data, indices) {
  d <- sample_n(data, 35) # allows boot to select sample
  attach(d)
  fit <- gnlr(num_seeds, dist="negative binomial", mu=formula, pmu=c(400,0.50), pshape=1)
  return(fit[[18]][2])
}

# bootstrapping with 1000 replications
set.seed(1)
results_alpha_ij_dry <- boot(data=subset_SEEDS, p2,
   R=50, formula=mu)

# view results
results_alpha_ij_dry
```

# ajj
```{r}
## Dry
## i = PLCO, j = VALO
##ajj
subset_SEEDS <- filter(SEEDS, focal_species == "VALO", background_species == "VALO" | background_species == "none", treatment == "Drought-stressed")

#m2_VALO_ajj_drought<-nls(num_seeds~lambda/(1+alpha*germination_j*background_density), data=subset_SEEDS, start=list(lambda=500, alpha=0.35))
#y<-subset_SEEDS$num_seeds
#x<-subset_SEEDS$background_density

#summary(m2_VALO_ajj_drought)
#confint(m2_VALO_ajj_drought)

attach(subset_SEEDS)
mu <- function(p) p[1]/(1+p[2]*background_density)
m2_VALO_ajj_drought <- gnlr(num_seeds, dist="negative binomial", mu=mu, pmu=c(500,0.5), pshape=1)

# Bootstrap 

# function to setimate parameters from given seed_count and background density
mu <- function(p) p[1]/(1+p[2]*background_density)

# function to obtain p1 from the data
p1 <- function(formula, data, indices) {
  d <- sample_n(data, 35) # allows boot to select sample
  attach(d)
  fit <- gnlr(num_seeds, dist="negative binomial", mu=formula, pmu=c(500,0.50), pshape=1)
  return(fit[[18]][1])
}

# bootstrapping with 1000 replications
set.seed(1)
results_lambda_j_dry <- boot(data=subset_SEEDS, p1,
   R=50, formula=mu)

# view results
results_lambda_j_dry

# function to obtain p2 from the data
p2 <- function(formula, data, indices) {
  d <- sample_n(data, 35) # allows boot to select sample
  attach(d)
  fit <- gnlr(num_seeds, dist="negative binomial", mu=formula, pmu=c(600,0.50), pshape=1)
  return(fit[[18]][2])
}

# bootstrapping with 1000 replications
set.seed(1)
results_alpha_jj_dry <- boot(data=subset_SEEDS, p2,
   R=50, formula=mu)

# view results
results_alpha_jj_dry
```

# aji
```{r}
## Dry
## i = PLCO, j = VALO
##ajj
subset_SEEDS <- filter(SEEDS, focal_species == "VALO", background_species == "PLCO" | background_species == "none", treatment == "Drought-stressed")

#m2_VALO_aji_drought<-nls(num_seeds~lambda/(1+alpha*germination_i*background_density), data=subset_SEEDS, start=list(lambda=500, alpha=0.35))
#y<-subset_SEEDS$num_seeds
#x<-subset_SEEDS$background_density

#summary(m2_VALO_aji_drought)
#confint(m2_VALO_aji_drought)

attach(subset_SEEDS)
mu <- function(p) p[1]/(1+p[2]*background_density)
m2_VALO_aji_drought <- gnlr(num_seeds, dist="negative binomial", mu=mu, pmu=c(636,0.40), pshape=1)

# Bootstrap 

# function to setimate parameters from given seed_count and background density
mu <- function(p) p[1]/(1+p[2]*background_density)


# function to obtain p2 from the data
p2 <- function(formula, data, indices) {
  d <- sample_n(data, 35) # allows boot to select sample
  attach(d)
  fit <- gnlr(num_seeds, dist="negative binomial", mu=formula, pmu=c(600, 0.50), pshape=1)
  return(fit[[18]][2])
}

# bootstrapping with 1000 replications
set.seed(1)
results_alpha_ji_dry <- boot(data=subset_SEEDS, p2,
   R=50, formula=mu)

# view results
results_alpha_ji_dry
```

# synthesize parameter estimates and calculate coexistence criterion 
```{r}
## Dry
## then niche overlap (p) = sqrt((aij/ajj)*(aji/aii))
## (kj/ki) = ((nj - 1)/(ni - 1))(sqrt((aij/ajj)*(aii/aji))
## (kj/ki) is here defined as the fitness ratio
## ni = (lamda_i*germination_i)/(1 - (1 - germination_i)(s_i))

(aij_drought = signif(m2_PLCO_aij_drought$coefficients[2], 4))
(ajj_drought = signif(m2_VALO_ajj_drought$coefficients[2], 4))
(aji_drought = signif(m2_VALO_aji_drought$coefficients[2], 4))
(aii_drought = signif(m2_PLCO_aii_drought$coefficients[2], 4))
(p_drought = sqrt((aij_drought/ajj_drought)*(aji_drought/aii_drought)))
(1 / p_drought)

(lambda_i_drought = (signif(m2_PLCO_aii_drought$coefficients[1], 5) + signif(m2_PLCO_aij_drought$coefficients[1], 5)) / 2)
(lambda_j_drought = (signif(m2_VALO_ajj_drought$coefficients[1], 5) + signif(m2_VALO_aji_drought$coefficients[1], 5)) / 2)

(ni_drought = (lambda_i_drought*germination_i)/(1 - (1 - germination_i)*(1)))
(nj_drought = (lambda_j_drought*germination_j)/(1 - (1 - germination_j)*(1)))
## demogrphic ratio (nj - 1/ ni - 1) "describes the degree to which species j produces more seeds (kjgj) per seed loss due to death or germination (1 - (1 - gj)(sj)) than species i" (godoy & levine 2014).
(demographic_ratio_drought <- (nj_drought - 1)/(ni_drought - 1))
## the ratio is > 1, thus j (valerianella) produces more seeds per seed loss giving it a higher demographic potential

## the competitive response ratio (sqrt((aij/ajj)*(aii/aji))) "describes the degree to which species j is more resistant to competition than species i. Notice that the numerator and denominator differ only in which species is responding to competition, and the arrangement of the interaction coefficients is not the same as in q." (Godoy & Levine 2014)
((aij_drought/ajj_drought)*(aii_drought/aji_drought))
(competitive_response_ratio_drought <- sqrt((aij_drought/ajj_drought)*(aii_drought/aji_drought)))
## greater than one means top numbers are larger, meaning i more sensitive to competition.
## less than one means bottom numbers are larger, meaning j more sensitive to competition.
## the ratio here is less than one, meaning i (Plectritis) is less sensitive to competition.
  
## the fitness ratio combines the demographic potential with the competitive response ratio
## "The demographic ratio and competitive response ratio therefore provide alternative pathways to competitive dominance, and their product, the fitness response ratio, predicts the winner of competition in the case of perfect niche overlap.
(fitness_ratio_drought = demographic_ratio_drought*competitive_response_ratio_drought)
## the fitness ratio is 0.656, meaning that Plectritis would win in the case of competition with complete niche overlap. 

## Coexistence requires both species to invade when rare (Chesson 2012). This condition is satisfied when (Godoy & Levine 2014): p < fitness ratio < 1/p
# p = 0.8364
# fitness_ratio = 0.771
# 1 / p = 1.196
ifelse(p_drought < fitness_ratio_drought & fitness_ratio_drought < 1/p_drought, "coexistence predicted", "coexistence not predicted")
```


#### Estimate parameters in wet treatment conditions
# aii
```{r}
## wet
## i = PLCO, j = VALO
##aii
subset_SEEDS <- filter(SEEDS, focal_species == "PLCO", background_species == "PLCO" | background_species == "none", treatment == "Well-watered")

attach(subset_SEEDS)
mu <- function(p) p[1]/(1+p[2]*background_density)
m2_PLCO_aii_wet <- gnlr(num_seeds, dist="negative binomial", mu=mu, pmu=c(400,0.40), pshape=1)  

# Bootstrap 

# function to setimate parameters from given seed_count and background density
mu <- function(p) p[1]/(1+p[2]*background_density)

# function to obtain p1 from the data
p1 <- function(formula, data, indices) {
  d <- sample_n(data, 35) # allows boot to select sample
  attach(d)
  fit <- gnlr(num_seeds, dist="negative binomial", mu=formula, pmu=c(400,0.50), pshape=1)
  return(fit[[18]][1])
}

# bootstrapping with 1000 replications
set.seed(1)
results_lambda_i_wet <- boot(data=subset_SEEDS, p1,
   R=50, formula=mu)

# view results
results_lambda_i_wet

# function to obtain p2 from the data
p2 <- function(formula, data, indices) {
  d <- sample_n(data, 35) # allows boot to select sample
  attach(d)
  fit <- gnlr(num_seeds, dist="negative binomial", mu=formula, pmu=c(400,0.50), pshape=1)
  return(fit[[18]][2])
}

# bootstrapping with 1000 replications
set.seed(1)
results_alpha_ii_wet <- boot(data=subset_SEEDS, p2,
   R=50, formula=mu)

# view results
results_alpha_ii_wet
```

# aij
```{r}
## Wet
## i = PLCO, j = VALO
##aij
subset_SEEDS <- filter(SEEDS, focal_species == "PLCO", background_species == "VALO" | background_species == "none", treatment == "Well-watered")

#m2_PLCO_aij_wet<-nls(num_seeds~lambda/(1+alpha*germination_j*background_density), data=subset_SEEDS, start=list(lambda=304, alpha=0.407))
#y<-subset_SEEDS$num_seeds
#x<-subset_SEEDS$background_density

#summary(m2_PLCO_aij_wet)
#confint(m2_PLCO_aij_wet)

attach(subset_SEEDS)
mu <- function(p) p[1]/(1+p[2]*background_density)
m2_PLCO_aij_wet <- gnlr(num_seeds, dist="negative binomial", mu=mu, pmu=c(400,0.40), pshape=1)  

# Bootstrap 

# function to setimate parameters from given seed_count and background density
mu <- function(p) p[1]/(1+p[2]*background_density)

# function to obtain p2 from the data
p2 <- function(formula, data, indices) {
  d <- sample_n(data, 35) # allows boot to select sample
  attach(d)
  fit <- gnlr(num_seeds, dist="negative binomial", mu=formula, pmu=c(400,0.50), pshape=1)
  return(fit[[18]][2])
}

# bootstrapping with 1000 replications
set.seed(1)
results_alpha_ij_wet <- boot(data=subset_SEEDS, p2,
   R=50, formula=mu)

# view results
results_alpha_ij_wet
```

# ajj
```{r}
## Wet
## i = PLCO, j = VALO
##ajj
subset_SEEDS <- filter(SEEDS, focal_species == "VALO", background_species == "VALO" | background_species == "none", treatment == "Well-watered")

#m2_VALO_ajj_wet<-nls(num_seeds~lambda/(1+alpha*germination_i*background_density), data=subset_SEEDS, start=list(lambda=96.424, alpha=0.088))
#y<-subset_SEEDS$num_seeds
#x<-subset_SEEDS$background_density

#summary(m2_VALO_ajj_wet)
#confint(m2_VALO_ajj_wet)

attach(subset_SEEDS)
mu <- function(p) p[1]/(1+p[2]*background_density)
m2_VALO_ajj_wet <- gnlr(num_seeds, dist="negative binomial", mu=mu, pmu=c(300,0.45), pshape=1) 

# Bootstrap 

# function to setimate parameters from given seed_count and background density
mu <- function(p) p[1]/(1+p[2]*background_density)

# function to obtain p1 from the data
p1 <- function(formula, data, indices) {
  d <- sample_n(data, 35) # allows boot to select sample
  attach(d)
  fit <- gnlr(num_seeds, dist="negative binomial", mu=formula, pmu=c(400,0.50), pshape=1)
  return(fit[[18]][1])
}

# bootstrapping with 1000 replications
set.seed(1)
results_lambda_j_wet <- boot(data=subset_SEEDS, p1,
   R=50, formula=mu)

# view results
results_lambda_j_wet

# function to obtain p2 from the data
p2 <- function(formula, data, indices) {
  d <- sample_n(data, 35) # allows boot to select sample
  attach(d)
  fit <- gnlr(num_seeds, dist="negative binomial", mu=formula, pmu=c(400,0.50), pshape=1)
  return(fit[[18]][2])
}

# bootstrapping with 1000 replications
set.seed(1)
results_alpha_jj_wet <- boot(data=subset_SEEDS, p2,
   R=50, formula=mu)

# view results
results_alpha_jj_wet
```

# aji
```{r}
## wet
## i = PLCO, j = VALO
##aji
subset_SEEDS <- filter(SEEDS, focal_species == "VALO", background_species == "PLCO" | background_species == "none", treatment == "Well-watered")

#m2_VALO_aji_wet<-nls(num_seeds~lambda/(1+alpha*germination_j*background_density), data=subset_SEEDS, start=list(lambda=615, alpha=0.63))
#y<-subset_SEEDS$num_seeds
#x<-subset_SEEDS$background_density

#summary(m2_VALO_aji_wet)
#confint(m2_VALO_aji_wet)

attach(subset_SEEDS)
mu <- function(p) p[1]/(1+p[2]*background_density)
m2_VALO_aji_wet <- gnlr(num_seeds, dist="negative binomial", mu=mu, pmu=c(325,0.38), pshape=1) 

# function to obtain p2 from the data
p2 <- function(formula, data, indices) {
  d <- sample_n(data, 35) # allows boot to select sample
  attach(d)
  fit <- gnlr(num_seeds, dist="negative binomial", mu=formula, pmu=c(400,0.50), pshape=1)
  return(fit[[18]][2])
}

# bootstrapping with 1000 replications
set.seed(1)
results_alpha_ji_wet <- boot(data=subset_SEEDS, p2,
   R=50, formula=mu)

# view results
results_alpha_ji_wet
```

# synthesize parameter estimates and calculate coexistence criterion 
```{r}
(aij_wet = signif(m2_PLCO_aij_wet$coefficients[2], 4))
(ajj_wet = signif(m2_VALO_ajj_wet$coefficients[2], 4))
(aji_wet = signif(m2_VALO_aji_wet$coefficients[2], 4))
(aii_wet = signif(m2_PLCO_aii_wet$coefficients[2], 4))

(p_wet = sqrt((aij_wet/ajj_wet)*(aji_wet/aii_wet)))
(1 / p_wet)

## (kj/ki) = ((nj - 1)/(ni - 1))(sqrt((aij/ajj)*(aii/aji))
## (kj/ki) is here defined as the fitness ratio
## ni = (lamda_i*germination_i)/(1 - (1 - germination_i)(s_i))
## currently don't have seed survival values (s_i), so will sub a value of 0.50 for both species.

(lambda_i_wet = (signif(m2_PLCO_aii_wet$coefficients[1], 5) + signif(m2_PLCO_aij_wet$coefficients[1], 5)) / 2)
(lambda_j_wet = (signif(m2_VALO_ajj_wet$coefficients[1], 5) + signif(m2_VALO_aji_wet$coefficients[1], 5)) / 2)

(ni_wet = (lambda_i_wet*germination_i)/(1 - (1 - germination_i)*(1)))
(nj_wet = (lambda_j_wet*germination_j)/(1 - (1 - germination_j)*(1)))
## demogrphic ratio (nj - 1/ ni - 1) "describes the degree to which species j produces more seeds (kjgj) per seed loss due to death or germination (1 - (1 - gj)(sj)) than species i" (godoy & levine 2014).
(demographic_ratio_wet <- (nj_wet - 1)/(ni_wet - 1))
## the ratio is < 1, thus i (Plectritis) produces more seeds per seed loss giving it a higher demographic potential

## the competitive response ratio (sqrt((aij/ajj)*(aii/aji))) "describes the degree to which species j is more resistant to competition than species i. Notice that the numerator and denominator differ only in which species is responding to competition, and the arrangement of the interaction coefficients is not the same as in q." (Godoy & Levine 2014)
((aij_wet/ajj_wet)*(aii_wet/aji_wet))
(competitive_response_ratio_wet <- sqrt((aij_wet/ajj_wet)*(aii_wet/aji_wet)))
## greater than one means top numbers are larger, meaning i more sensitive to competition.
## less than one means bottom numbers are larger, meaning j more sensitive to competition.
## the ratio here is more than one, meaning j (Valerianella) is less sensitive to competition.

## the fitness ratio combines the demographic potential with the competitive response ratio
## "The demographic ratio and competitive response ratio therefore provide alternative pathways to competitive dominance, and their product, the fitness response ratio, predicts the winner of competition in the case of perfect niche overlap.
(fitness_ratio_wet = demographic_ratio_wet*competitive_response_ratio_wet)
## the fitness ratio is 0.995, meaning that i (Plectritis) would win in the case of competition with complete niche overlap. 

## Coexistence requires both species to invade when rare (Chesson 2012). This condition is satisfied when (Godoy & Levine 2014): p (niche overlap) < fitness ratio < 1/p
# p = 0.8012
# fitness_ratio = 0.995
# 1 / p = 1.2482
ifelse(p_wet < fitness_ratio_wet & fitness_ratio_wet < 1/p_wet, "coexistence predicted", "coexistence not predicted")
```

## Estimate parameter error using propogate (dry conditions)
```{r}
# enter parameter estimates with sd's
lambda_p <- c(365.9, 17.7)
lambda_v <- c(650, 57.4)
alpha_pv <- c(.29, 0.03)
alpha_vp <- c(0.45, 0.05)
alpha_pp <- c(0.25, 0.02)
alpha_vv <- c(0.73, 0.11)

# enter data from germination test
germination_p = c(0.670, 0)
germination_v = c(0.891, 0)
survival_p = c(1, 0)
survival_v = c(1, 0)

## calculate niche overlap
# define expression
EXPR1 <- expression(sqrt((alpha_pv/alpha_vv)*
                           (alpha_vp/alpha_pp)))
# make a vector of pertinent parameter estimates
DF1 <- cbind(alpha_pv,
             alpha_vp, alpha_pp, alpha_vv)
# run propogate and display results
(RES1 <- propagate(expr = EXPR1, data = DF1, type = "stat",
                  do.sim = TRUE, verbose = TRUE,
                  nsim = 1000000))

## calculate competitive response ratio
EXPR2 <- expression(sqrt((alpha_pv/alpha_vv)*(alpha_pp/alpha_vp)))
DF2 <- cbind(alpha_pv,
             alpha_vp, alpha_pp, alpha_vv)
RES2 <- propagate(expr = EXPR2, data = DF2, type = "stat",
                  do.sim = TRUE, verbose = TRUE,
                  nsim = 1000000)
RES2

## calculate demographic ratio
EXPR3 <- expression(
  ((lambda_v*germination_v)/(germination_v))/
  ((lambda_p*germination_p)/(germination_p))
)
DF3 <- cbind(lambda_p, lambda_v, germination_p, germination_v)
RES3 <- propagate(expr = EXPR3, data = DF3, type = "stat",
                  do.sim = TRUE, verbose = TRUE,
                  nsim = 1000000)
RES3

## calculate fitness ratio
# estimated this mean directly from the alpha and lambda parameters
# rather than multiplying the demographic and competitive response ratios above
EXPR4 <- expression((sqrt((alpha_pv/alpha_vv)*(alpha_pp/alpha_vp)))*
                      (
                        ((lambda_v*germination_v)/(germination_v))/
                          ((lambda_p*germination_p)/(germination_p))
                      )
)
DF4 <- cbind(lambda_p, lambda_v, germination_p, germination_v, 
             alpha_pv, alpha_vp, alpha_pp, alpha_vv)
RES4 <- propagate(expr = EXPR4, data = DF4, type = "stat",
                  do.sim = TRUE, verbose = TRUE,
                  nsim = 1000000)
RES4
```

## Estimate parameter error using propogate (wet conditions)
```{r}
lambda_p <- c(391.23, 27.8)
lambda_v <- c(327.4, 24.8)
alpha_pv <- c(0.27, 0.09)
alpha_vp <- c(0.42, 0.04)
alpha_pp <- c(0.37, 0.06)
alpha_vv <- c(0.47, 0.06)

germination_p = c(0.670, 0)
germination_v = c(0.891, 0)
survival_p = c(1, 0)
survival_v = c(1, 0)

## niche overlap
EXPR1_wet <- expression(sqrt((alpha_pv/alpha_vv)*
                           (alpha_vp/alpha_pp)))
DF1_wet <- cbind(alpha_pv,
             alpha_vp, alpha_pp, alpha_vv)
RES1_wet <- propagate(expr = EXPR1_wet, data = DF1_wet, type = "stat",
                  do.sim = TRUE, verbose = TRUE,
                  nsim = 1000000)
RES1_wet

# competitive response ratio
EXPR2_wet <- expression(sqrt((alpha_pv/alpha_vv)*(alpha_pp/alpha_vp)))
DF2_wet <- cbind(alpha_pv,
             alpha_vp, alpha_pp, alpha_vv)

RES2_wet <- propagate(expr = EXPR2_wet, data = DF2_wet, type = "stat",
                  do.sim = TRUE, verbose = TRUE,
                  nsim = 1000000)
RES2_wet


# demographic ratio
EXPR3_wet <- expression(
  ((lambda_v*germination_v)/(germination_v)) / ((lambda_p*germination_p)/(germination_p))
)

DF3_wet <- cbind(lambda_p, lambda_v, germination_p, germination_v)

RES3_wet <- propagate(expr = EXPR3_wet, data = DF3_wet, type = "stat",
                  do.sim = TRUE, verbose = TRUE,
                  nsim = 1000000)
RES3_wet


## calculate fitness ratio
# estimated this mean directly from the alpha and lambda parameters
# rather than multiplying the demographic and competitive response ratios above

EXPR4_wet <- expression((sqrt((alpha_pv/alpha_vv)*(alpha_pp/alpha_vp)))*
                      (
                        ((lambda_v*germination_v)/(germination_v))/
                          ((lambda_p*germination_p)/(germination_p))
                      )
)
DF4_wet <- cbind(lambda_p, lambda_v, germination_p, germination_v, 
             alpha_pv, alpha_vp, alpha_pp, alpha_vv)
RES4_wet <- propagate(expr = EXPR4_wet, data = DF4_wet, type = "stat",
                  do.sim = TRUE, verbose = TRUE,
                  nsim = 1000000)
RES4_wet
```

## Plot Coexistence parameters (Fig 1A)
```{r}
#############################
#############################
df <- read.csv("parameter_means_sd.csv")


fun.1 <- function(x) 1 - x
fun.2 <- function(x) 1 + x


T <- ggplot(df, aes(x=(1-mean_p), y=mean_k, colour=X)) + 
  geom_errorbar(aes(ymin=mean_k-se_k, ymax=mean_k+se_k), width=.05, size = 1) +
  geom_point(size = 6) 
T <- T + geom_errorbarh(aes(xmin=(1-mean_p-se_p), xmax=(1-mean_p+se_p), height=.05)) 

T <- T + theme_bw() + 
  theme(plot.title = element_text(hjust = 0), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank()) + 
  theme(legend.title=element_text()) + labs(x=substitute( paste("niche difference (1 - ",  italic('p'), ")" )), y="average fitness ratio") +
  theme(legend.position="none")
# color black for dry, grey for wet
T <- T + scale_color_manual(values=c("black", "grey"))
T <- T + theme(axis.line = element_line(size = 2))
T <- T + theme(axis.title.x = element_text(size = 20),
               # Y axis title
               axis.title.y = element_text(size = 20),
               # X axis text
               axis.text.x = element_text(
                 size = 16,
                 angle = 0,
                 vjust = 1
               ),
               # Y axis text
               axis.text.y = element_text(size = 16))
T <- T + theme(plot.title = element_text(size = 20, face = "bold"))
T <- T + scale_y_continuous(labels = function(x) round(as.numeric(x), digits=1), expand = c(0, 0), limits = c(0, 2), breaks = c(0, .5, 1, 1.5, 2.0))
T <- T + scale_x_continuous(labels = function(x) round(as.numeric(x), digits=1), expand = c(0, 0), limits = c(0, 1.05), breaks = c(0, .2, .4, .6, .8, 1))
T <- T + geom_hline(yintercept= 1, linetype="dashed", 
                    color = "black", size=1)

T <- T + stat_function(fun = fun.1, size = 1.5, colour = "grey30")
T <- T + stat_function(fun = fun.2, size = 1.5, colour = "grey30")


grob <- grobTree(textGrob(substitute(paste(italic("V. locusta "), "dominant")), x=0.05,  y=0.83, hjust=0,
                          gp=gpar(col="black", fontsize=20)))
grob2 <- grobTree(textGrob(substitute(paste(italic("P. congesta "), "dominant")), x=0.05,  y=0.05, hjust=0,
                          gp=gpar(col="black", fontsize=20)))
grob3 <- grobTree(textGrob("exclusion", x=0.4,  y=0.95, hjust=0,
                          gp=gpar(col="black", fontsize=20)))
grob4 <- grobTree(textGrob("coexistence", x=0.65,  y=0.6, hjust=0,
                          gp=gpar(col="black", fontsize=20)))
grob5 <- grobTree(textGrob("exclusion", x=0.4,  y=0.15, hjust=0,
                           gp=gpar(col="black", fontsize=20)))
grob6 <- grobTree(textGrob("coexistence", x=0.65,  y=0.4, hjust=0,
                           gp=gpar(col="black", fontsize=20)))


T <- T + annotation_custom(grob)
T <- T + annotation_custom(grob2)
T <- T + annotation_custom(grob3)
T <- T + annotation_custom(grob4)
T <- T + annotation_custom(grob5)
T <- T + annotation_custom(grob6)
T <- T + coord_fixed(ratio = 0.5)
T <- T + theme(axis.line = element_line(colour = 'black', size = 1))
T
```

## Calculare low-density invasion growth rates (and plot Fig 1B)
```{r}
# invasion growth rates

# wet conditions
# species j = Valerianella  
# species i = Plectritis

lambda_p <- c(391.23, 27.8)
lambda_v <- c(327.4, 24.8)
alpha_pv <- c(0.27, 0.09)
alpha_vp <- c(0.42, 0.04)
alpha_pp <- c(0.37, 0.06)
alpha_vv <- c(0.47, 0.06)

## From Hart et al. 2019
## "we quantified the invasion growth rate of sic (j) invading sic (i).
## using the following equation": 
## j_inv <- lambdaj / (1 + aji*((lambdai - 1)/aii))
## "where j is at vanishingly small density and
## N(i),t has been replaced by the expression
## which quantifies the equilibrium population
## density of i." 

## and likewise for species i
## i_inv <- lambdai / (1 + aij*((lambdaj - 1)/ajj))
## 

## invasion growth rate Valerianella (wet)
EXPR1 <- expression(lambda_v / (1 + alpha_vp*((lambda_p - 1)/alpha_pp)))
DF1 <- cbind(lambda_v, lambda_p,
             alpha_vp, alpha_pp)
RES1 <- propagate(expr = EXPR1, data = DF1, type = "stat",
                  do.sim = TRUE, verbose = TRUE,
                  nsim = 1000000)
RES1

## invasion growth rate Plectritis (wet)
EXPR2 <- expression(lambda_p / (1 + alpha_pv*((lambda_v - 1)/alpha_vv)))
DF2 <- cbind(lambda_v, lambda_p,
             alpha_pv, alpha_vv)
RES2 <- propagate(expr = EXPR2, data = DF2, type = "stat",
                  do.sim = TRUE, verbose = TRUE,
                  nsim = 1000000)
RES2

# dry conditions
# species j = Valerianella  
# species i = Plectritis

# enter parameter estimates with sd's
lambda_p_dry <- c(365.9, 17.7)
lambda_v_dry <- c(650, 57.4)
alpha_pv_dry <- c(.29, 0.03)
alpha_vp_dry <- c(0.45, 0.05)
alpha_pp_dry <- c(0.25, 0.02)
alpha_vv_dry <- c(0.73, 0.11)


## invasion growth rate Valerianella (dry)
EXPR1_dry <- expression(lambda_v_dry / (1 + alpha_vp_dry*((lambda_p_dry - 1)/alpha_pp_dry)))
DF1_dry <- cbind(lambda_v_dry, lambda_p_dry,
             alpha_vp_dry, alpha_pp_dry)
RES1_dry <- propagate(expr = EXPR1_dry, data = DF1_dry, type = "stat",
                  do.sim = TRUE, verbose = TRUE,
                  nsim = 1000000)
RES1_dry

## invasion growth rate Plectritis (dry)
EXPR2_dry <- expression(lambda_p_dry / (1 + alpha_pv_dry*((lambda_v_dry - 1)/alpha_vv_dry)))
DF2_dry <- cbind(lambda_v_dry, lambda_p_dry,
             alpha_pv_dry, alpha_vv_dry)
RES2_dry <- propagate(expr = EXPR2_dry, data = DF2_dry, type = "stat",
                  do.sim = TRUE, verbose = TRUE,
                  nsim = 1000000)
RES2_dry


mean_inv_growth_rate_j <- as.data.frame(rbind(0.74, 0.99))
sd_inv_growth_rate_j <- as.data.frame(rbind(0.15, 0.17))
mean_inv_growth_rate_i <- as.data.frame(rbind(2.08, 1.43))
sd_inv_growth_rate_i <- as.data.frame(rbind(0.77, 0.29))

colnames(mean_inv_growth_rate_j) <- "mean_inv_growth_rate"
colnames(sd_inv_growth_rate_j) <- "sd_inv_growth_rate"
colnames(mean_inv_growth_rate_i) <- "mean_inv_growth_rate"
colnames(sd_inv_growth_rate_i) <- "sd_inv_growth_rate"

mean_inv_growth_rate_j$treatment <- rbind("wet", "dry")
sd_inv_growth_rate_j$treatment <- rbind("wet", "dry")
mean_inv_growth_rate_i$treatment <- rbind("wet", "dry")
sd_inv_growth_rate_i$treatment <- rbind("wet", "dry")

mean_inv_growth_rate_j$species <- rbind("Valerianella", "Valerianella")
sd_inv_growth_rate_j$species <- rbind("Valerianella", "Valerianella")
mean_inv_growth_rate_i$species <- rbind("Plectritis", "Plectritis")
sd_inv_growth_rate_i$species <- rbind("Plectritis", "Plectritis")

Valerianella <- merge(mean_inv_growth_rate_j, sd_inv_growth_rate_j)
Plectritis <- merge(mean_inv_growth_rate_i, sd_inv_growth_rate_i)

df2 <- rbind(Valerianella, Plectritis)
df2$X <- c("1", "2", "3", "4")


#############################
# plot invasion growth rates

E <- ggplot(df2, aes(x=species, y=mean_inv_growth_rate, colour=treatment)) + 
  geom_errorbar(aes(ymin=mean_inv_growth_rate-sd_inv_growth_rate, 
                    ymax=mean_inv_growth_rate+sd_inv_growth_rate), 
                width=.2, size = 1, position=position_dodge(width = 0.5)) 
E <- E + geom_point(size = 6, position=position_dodge(width=0.5)) 
E <- E + theme_bw() 
E <- E + theme(plot.title = element_text(hjust = 0), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank()) + 
  theme(legend.title=element_text()) + labs(x="treatment", y="invasion growth rate") 
E <- E + theme(legend.title=element_text(size=20), legend.text=element_text(size=20))
E <- E + scale_color_manual(values=c("black", "grey"))
E <- E + theme(axis.line = element_line(size = 2))
E <- E + theme(axis.title.x = element_blank(),
               # Y axis title
               axis.title.y = element_text(size = 20),
               # X axis text
               axis.text.x = element_text(
                 size = 20,
                 angle = 0,
                 vjust = .5,
                 face = "italic"
               ),
               # Y axis text
               axis.text.y = element_text(size = 14))
E <- E + theme(plot.title = element_text(size = 20, face = "bold"))
E <- E + scale_y_continuous(expand = c(0, 0), limits = c(0, 4), breaks = c(0, 1, 2, 3, 4))
E <- E + geom_hline(yintercept= 1, linetype="dashed", 
                    color = "black", size=1)

E <- E + theme(legend.position = c(0.75, 0.9))
E <- E + coord_fixed(ratio = 0.45)
E <- E + theme(axis.line = element_line(colour = 'black', size = 1))
E <- E + scale_x_discrete(labels = c('P. congesta','V. locusta'))
E
```

## Plot Fig 1
```{r}

ggarrange(T + theme(plot.margin = margin(r = 0, l = 0, t = 45, b = 25)), 
          E + theme(plot.margin = margin(r = 0, l = 20, t = 0, b = 0)), 
          labels = c("(a)", "(b)"), font.label = list(size = 20))

```

## Plot supplementary figure S2
```{r}
## Plot drought
subset_SEEDS_drought <- filter(SEEDS, treatment == "Drought-stressed")
subset_SEEDS_drought_PLCO <- filter(SEEDS, treatment == "Drought-stressed", focal_species == "PLCO")
subset_SEEDS_drought_VALO <- filter(SEEDS, treatment == "Drought-stressed", focal_species == "VALO")
subset_SEEDS_aii <- filter(SEEDS, focal_species == "PLCO", background_species == "PLCO", treatment == "Drought-stressed")
subset_SEEDS_aij <- filter(SEEDS, focal_species == "PLCO", background_species == "VALO", treatment == "Drought-stressed")
subset_SEEDS_aji <- filter(SEEDS, focal_species == "VALO", background_species == "PLCO", treatment == "Drought-stressed")
subset_SEEDS_ajj <- filter(SEEDS, focal_species == "VALO", background_species == "VALO", treatment == "Drought-stressed")



####### (A) plot PLCO focals in dry treatment
P <- ggplot(subset_SEEDS_drought_PLCO, aes(x = background_density, y = num_seeds, color = interaction(focal_species, background_species, sep=" vs. ",lex.order=TRUE)))
P <- P + geom_jitter(aes(background_density, num_seeds, shape=background_species), width = 1, size = 2, alpha=.7) 
# function = number_seeds ~ lambda / 1 + alpha*background_density
function_1 <- function(x) 366/(1+.25*x)
P <- P + stat_function(data = subset_SEEDS_aii, fun = function_1, size = 1, colour = "black")
function_2 <- function(x) 366/(1+.29*x)
P <- P + stat_function(data = subset_SEEDS_aij, fun = function_2, size = 1, colour = "black", linetype = "dashed")
P <- P + scale_shape_manual(values=c(19,1))
P <- P + theme_bw() + theme(plot.title = element_text(hjust = 0), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank()) + theme(legend.title=element_blank()) + labs(x="Neighbor density", y="P. congesta seeds/plant")
P <- P + theme(legend.position="none")
P <- P + scale_color_manual(values=c("black", "black"))
P <- P + theme(axis.line = element_line(size = 2))
P <- P + theme(axis.title.x = element_text(size = 20),
               # Y axis title
               axis.title.y = element_text(size = 20),
               # X axis text
               axis.text.x = element_text(
                 size = 16,
                 angle = 0,
                 vjust = 1
               ),
               # Y axis text
               axis.text.y = element_text(size = 16))
P <- P + theme(axis.line = element_line(size = 1))
P

#### (B) Plot VALO focals in dry
R <- ggplot(subset_SEEDS_drought_VALO, aes(x = background_density, y = num_seeds, color = interaction(focal_species, background_species, sep=" vs. ",lex.order=TRUE)))
R <- R + geom_jitter(aes(background_density, num_seeds, shape=background_species), width = 1, size = 2, alpha=.7) 
function_3 <- function(x) 650 /(1+.45*x)
R <- R + stat_function(data = subset_SEEDS_aji, fun = function_3, size = 1, colour = "black", linetype = "dashed")
function_4 <- function(x) 650 /(1+.73*x)
R <- R + stat_function(data = subset_SEEDS_ajj, fun = function_4, size = 1, colour = "black")
R <- R + scale_shape_manual(values=c(1,19))
R <- R + theme_bw() + theme(plot.title = element_text(hjust = 0), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank()) + theme(legend.title=element_blank()) + labs(x="Neighbor Density", y="V. locusta seeds/plant")
R <- R + theme(legend.position="none")
R <- R + scale_color_manual(values=c("black", "black"))
R <- R + theme(axis.line = element_line(size = 2))
R <- R + theme(axis.title.x = element_text(size = 20),
               # Y axis title
               axis.title.y = element_text(size = 20),
               # X axis text
               axis.text.x = element_text(
                 size = 16,
                 angle = 0,
                 vjust = 1
               ),
               # Y axis text
               axis.text.y = element_text(size = 16))
R <- R + theme(axis.line = element_line(size = 1))
R

## Plot wet
subset_SEEDS_wet <- filter(SEEDS, treatment == "Well-watered")
subset_SEEDS_wet_PLCO <- filter(SEEDS, treatment == "Well-watered", focal_species == "PLCO")
subset_SEEDS_wet_VALO <- filter(SEEDS, treatment == "Well-watered", focal_species == "VALO")
subset_SEEDS_aii_wet <- filter(SEEDS, focal_species == "PLCO", background_species == "PLCO", treatment == "Well-watered")
subset_SEEDS_aij_wet <- filter(SEEDS, focal_species == "PLCO", background_species == "VALO", treatment == "Well-watered")
subset_SEEDS_aji_wet <- filter(SEEDS, focal_species == "VALO", background_species == "PLCO", treatment == "Well-watered")
subset_SEEDS_ajj_wet <- filter(SEEDS, focal_species == "VALO", background_species == "VALO", treatment == "Well-watered")

##### Plot PLCO focals in wet treatment
Q <- ggplot(subset_SEEDS_wet_PLCO, aes(x = background_density, y = num_seeds, color = interaction(focal_species, background_species, sep=" vs. ",lex.order=TRUE)))
Q <- Q + geom_jitter(aes(background_density, num_seeds, shape = background_species), width = 1, size = 2, alpha = .7) 
function_5 <- function(x) 391 /(1+.37*x)
Q <- Q + stat_function(data = subset_SEEDS_aii_wet, fun = function_5, size = 1, colour = "black")
function_6 <- function(x) 391 /(1+.27*x)
Q <- Q + stat_function(data = subset_SEEDS_aij_wet, fun = function_6, size = 1, colour = "black", linetype = "dashed")
Q <- Q + scale_shape_manual(values=c(19,1))
Q <- Q + theme_bw() + theme(plot.title = element_text(hjust = 0), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank()) + theme(legend.title=element_blank()) + labs(x="Neighbor Density", y="P. congesta seeds/plant")
Q <- Q + theme(legend.position="none")
Q <- Q + scale_color_manual(values=c("black", "black"))
Q <- Q + theme(axis.line = element_line(size = 2))
Q <- Q + theme(axis.title.x = element_text(size = 20),
               # Y axis title
               axis.title.y = element_text(size = 20),
               # X axis text
               axis.text.x = element_text(
                 size = 16,
                 angle = 0,
                 vjust = 1
               ),
               # Y axis text
               axis.text.y = element_text(size = 16))
Q <- Q + theme(axis.line = element_line(size = 1))
Q

##### plot VALO focals in wet
S <- ggplot(subset_SEEDS_wet_VALO, aes(x = background_density, y = num_seeds, color = interaction(focal_species, background_species, sep=" vs. ",lex.order=TRUE)))
S <- S + geom_jitter(aes(background_density, num_seeds, shape = background_species), width = 1, size = 2, alpha = .7) 
function_7 <- function(x) 327 /(1+.47*x)
S <- S + stat_function(data = subset_SEEDS_aji_wet, fun = function_7, size = 1, colour = "black", linetype = "dashed")
function_8 <- function(x) 327 /(1+.42*x)
S <- S + stat_function(data = subset_SEEDS_ajj_wet, fun = function_8, size = 1, colour = "black")
S <- S + scale_shape_manual(values=c(1,19))
S <- S + theme_bw() + theme(plot.title = element_text(hjust = 0), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), panel.border = element_blank()) + theme(legend.title=element_blank()) + labs(x="Neighbor Density", y="V. locusta seeds/plant")
S <- S + theme(legend.position="none")
S <- S + scale_color_manual(values=c("black", "black"))
S <- S + theme(axis.line = element_line(size = 2))
S <- S + theme(axis.title.x = element_text(size = 20),
               # Y axis title
               axis.title.y = element_text(size = 20),
               # X axis text
               axis.text.x = element_text(
                 size = 16,
                 angle = 0,
                 vjust = 1
               ),
               # Y axis text
               axis.text.y = element_text(size = 16))
S <- S + theme(axis.line = element_line(size = 1))
S



library(cowplot)
legend_b <- get_legend(P + theme(legend.position="bottom"))
require(grid)
require(gridExtra)
grid.arrange(P, R, Q, S,  ncol=2, nrow=2, 
             top=textGrob("", gp=gpar(fontsize=20,font=7)))
ggarrange(P + theme(plot.margin = margin(r = 0, l = 20, t = 20, b = 0)), 
          Q + theme(plot.margin = margin(r = 0, l = 20, t = 20, b = 0)), 
          R + theme(plot.margin = margin(r = 0, l = 20, t = 20, b = 0)), 
          S + theme(plot.margin = margin(r = 0, l = 20, t = 20, b = 0)),
          labels = c("(a)", "(b)", "(c)", "(d)"), font.label = list(size = 20))
```